Gringo 2.0

### 2008-07-30 doc/documentation/paper.tex
Further bugfixes in the documentation, changed some error messages and one typo
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@414 3e02f596-3131-0410-8352-e04eeaec37cc

\paragraph{\texttt{--convert}}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{\texttt{-l}}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explaned in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{\texttt{-c}}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answersets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore adouble minus ``\texttt{--}'' must be given after all gringo related parameters are set. After them die clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output hold as clasps internal interface currently reads smodels format.
\paragraph{\texttt{-p}}
Parameter ``\texttt{-p}'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo.
\paragraph{\texttt{-g [1-7]}}
This parameter invokes the new output format for the intermediate languages ASPils \cite{atleastthisreferenceishouldgive}. The optional number can range from $1$ to $7$ and refers to one of the following normalforms (Simple=1, SModels=2, DisjunctiveLogicProgram=3, CModels=4, CModelsExtended=5, DLV=6, Conglomeration=7). If no normalform is specified, conglomeration(7)  is assumed.

@@

\subsection{Meta-Syntax}
The ``\#'' sign in front of all meta expressions is optional and can therefore be omitted.
\paragraph{hide}
The hide construct is used to hide atoms from the output of the ASP solver.
Two different syntaxes are supported by gringo.
The first 
\begin{verbatim}
 #hide name/arity.
\end{verbatim}
and the second
\begin{verbatim}
 #hide name(X1, ... , Xn).
\end{verbatim}
Both variants hide the predicate ``name''.
The arity of the predicate can be given explicit or implicit with a list of variables.
The variables are just placeholder and do not unify with other variables. For example hide p(X,X) would also hide the fact p(a,b). The second syntax is implemented for compatibility reasons only.
If the predicate that has to be hidden has arity zero the syntax is
\begin{verbatim}
 #hide name.
\end{verbatim}
Not to give explicitly all predicates that are hidden, only the ones which are visible the statement
\begin{verbatim}
 #hide.
\end{verbatim}
in conjunction with the ``show'' statement is used.
The ``show'' statement has the same syntax as ``hide'' but uses the keyword ``show''.
\begin{verbatim}
 #hide.
 #show p/3.
\end{verbatim}
The example hides all atoms except the predicate $p$ with arity $3$.
\paragraph{const}
Within the ``const'' statement, a default value for constants can be defined in the logic program.
while they are normally defined via the command line options.
A default value for a constant $n$ can be easily be set.
\begin{verbatim}
 #const n=5.
\end{verbatim}
\paragraph{domain}
A domain statement states that a specific variable has a global domain restriction.
The statement
\begin{verbatim}
 #domain time(T).
\end{verbatim}
makes sure that the variable ``T'' is always bound to the domain of ``time''.
This has the same effect as adding the domain ``time(T)'' to every rule the variable ``T'' occurs in.
\paragraph{compute}
The compute statement is a hint for the solver which models to compute.
Is has the syntax
\begin{verbatim}
 #compute number {a1, ..., aN, not aN+1, ..., not aM}.
\end{verbatim}
$number$ is the number of models that shall be computed by the answerset solver and can be omitted. If it is zero, no restriction to the number of models is applied.
All positive atoms in the compute statement must be in the computed models, and all negative atoms do not have to be in the models of the program.
The compute statement is solver dependent and supported by every solver.
\paragraph{Comments}
Comments included in the input are line based.
Every line starting with ``\%'' is a comment. The whole line is omitted and not to be handled by the grounder.

@@

\begin{itemize}
 \item 
\begin{verbatim}
 "Warning: sum with negative bounds in the body of a rule"
\end{verbatim}
This warning occurs if a sum aggregate with negative weights has been found and the output format smodels is used or the output is redirected to clasp. In this format negative bounds are not allowed.
 \item 
\begin{verbatim}
 "Warning: multiple definitions of #const n"
\end{verbatim}
This warning is printed if more than one \#const statement is used to set the same constant.
The last attached value is used for grounding.
\item
\begin{verbatim}
 "Warning: p/3 is never defined"
\end{verbatim}
This warnings states that an atom is used in the body of a rule that never occurs in the head.
``p'' is the name of the predicate and ``3'' is the arity.
\item
\begin{verbatim}
 "got ground program i hope u have enough memory :)"
\end{verbatim}
Use the command line option ``--convert'' to convert a ground program into another format.

@@

\subsection{Errors}
gringo2.0 stops the execution and prints an error message to standard error output.
\begin{itemize}
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  non domain predicates : { ... }"
\end{verbatim}
This error message gives a list of predicates that are not domain restricted.
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  weakly restricted variables: { ... }"
\end{verbatim}
This error message gives a list of predicates that are not lambda restricted. 
\item
\begin{verbatim}
 "syntax error on line Y column Z unexpected token: X"
\end{verbatim}
The parser can not read the input format.
\begin{verbatim}
 "Disjunction not allowed on this normalform."
 "Count aggregate/cardinality constraint not allowed
  in this normalform."
 "Count aggregate/cardinality constraint not allowed 
  with non trivial upper bound in this normalform."
 "Count aggregate/cardinality constraint not allowed 
  with negative lower bound in this normalform."
 "Sum aggregate/weight constraint not allowed
  in this normalform."
 "Sum aggregate/weight constraint not allowed 
  with non trivial upper bound in this normalform."
 "Sum aggregate/weight constraint not allowed 
  with negative lower bound in this normalform."
 "Sum aggregate/weight constraint not allowed
  with negative weights in this normalform."
 "Sum aggregate/weight constraint not allowed
  with negative weights in this normalform."
 "Max aggregate not allowed in this normalform."
 "Max aggregate not allowed
  with negative weights in this normalform."
 "Min aggregate not allowed in this normalform."
 "Min aggregate not allowed 
  with negative weights in this normalform."
 "Optimize statement not allowed in this normalform."
\end{verbatim}
%\begin{verbatim}
% "Times aggregate not allowed in this normalform."
% "Times aggregate not allowed
%  with negative weights in this normalform."
%\end{verbatim}
These errors belong to the ASPils output format. Another normal form as to be chosen.
\end{itemize}

>>>>

\paragraph{\texttt{--convert}}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{\texttt{-l}}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explaned in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{\texttt{-c}}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answersets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore adouble minus ``\texttt{--}'' must be given after all gringo related parameters are set. After them die clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output hold as clasps internal interface currently reads smodels format.
\paragraph{\texttt{-p}}
Parameter ``\texttt{-p}'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo.
\paragraph{\texttt{-g [1-7]}}
This parameter invokes the new output format for the intermediate languages ASPils \cite{atleastthisreferenceishouldgive}. The optional number can range from $1$ to $7$ and refers to one of the following normal forms (Simple=1, SModels=2, DisjunctiveLogicProgram=3, CModels=4, CModelsExtended=5, DLV=6, Conglomeration=7). If no normal form is specified, conglomeration(7)  is assumed.

@@

\subsection{Meta-Syntax}
The ``\#'' sign in front of all meta expressions is optional but is it strongly recommend to use it.
It separates the meta directives from the logic program more clearly and can prevent severe mistakes,
 confusing those constructs with atoms as shown in the example.
\begin{verbatim}
 hide.
 hidden :- hide.
\end{verbatim}
This program hides all predicates from the output of the ASP solver instead of deriving the facts \texttt{hide} and \texttt{hidden}.

\paragraph{hide}
The hide construct is used to hide atoms from the output of the ASP solver.
Two different syntaxes are supported by gringo.
The first 
\begin{verbatim}
 #hide name/arity.
\end{verbatim}
and the second
\begin{verbatim}
 #hide name(X1, ... , Xn).
\end{verbatim}
Both variants hide the predicate ``name''.
The arity of the predicate can be given explicit or implicit with a list of variables.
The variables are just placeholder and do not unify with other variables. For example hide p(X,X) would also hide the fact p(a,b). The second syntax is implemented for compatibility reasons only.
If the predicate that has to be hidden has arity zero the syntax is
\begin{verbatim}
 #hide name.
\end{verbatim}
Not to give explicitly all predicates that are hidden, only the ones which are visible the statement
\begin{verbatim}
 #hide.
\end{verbatim}
in conjunction with the ``show'' statement is used.
The ``show'' statement has the same syntax as ``hide'' but uses the keyword ``show''.
\begin{verbatim}
 #hide.
 #show p/3.
\end{verbatim}
The example hides all atoms except the predicate $p$ with arity $3$.
\paragraph{const}
Within the ``const'' statement, a default value for constants can be defined in the logic program.
while they are normally defined via the command line options.
A default value for a constant $n$ can be easily be set.
\begin{verbatim}
 #const n=5.
\end{verbatim}
\paragraph{domain}
A domain statement states that a specific variable has a global domain restriction.
The statement
\begin{verbatim}
 #domain time(T).
\end{verbatim}
makes sure that the variable ``T'' is always bound to the domain of ``time''.
This has the same effect as adding the domain ``time(T)'' to every rule the variable ``T'' occurs in.
\paragraph{compute}
The compute statement is a hint for the solver which models to compute.
Is has the syntax
\begin{verbatim}
 #compute number {a1, ..., aN, not aN+1, ..., not aM}.
\end{verbatim}
$number$ is the number of models that shall be computed by the answerset solver and can be omitted. If it is zero, no restriction to the number of models is applied.
All positive atoms in the compute statement must be in the computed models, and all negative atoms do not have to be in the models of the program.
The compute statement is solver dependent and supported by every solver.
\paragraph{Comments}
Comments included in the input are line based.
Every line starting with ``\%'' is a comment. The whole line is omitted and not to be handled by the grounder.

@@

\begin{itemize}
 \item 
\begin{verbatim}
 "Warning: sum with negative bounds in the body of a rule"
\end{verbatim}
This warning occurs if a sum aggregate with negative weights has been found and the output format smodels is used or the output is redirected to clasp. In this format negative bounds are compiled away using the transformation described in \cite{lparseManual}. This may lead to counterintuitive results.
 \item 
\begin{verbatim}
 "Warning: multiple definitions of #const n"
\end{verbatim}
This warning is printed if more than one \#const statement is used to set the same constant.
The last attached value is used for grounding.
\item
\begin{verbatim}
 "Warning: p/3 is never defined"
\end{verbatim}
This warnings states that an atom is used in the body of a rule that never occurs in the head.
``p'' is the name of the predicate and ``3'' is the arity.
\item
\begin{verbatim}
 "got ground program i hope u have enough memory :)"
\end{verbatim}
Use the command line option ``--convert'' to convert a ground program into another format.
For converting already ground programs to an output language a special module has been implemented that uses far less memory. Because trying to ground an already variable free program could need a huge amount of memory.

@@

\subsection{Errors}
gringo2.0 stops the execution and prints an error message to standard error output.
\begin{itemize}
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  non domain predicates : { ... }"
\end{verbatim}
This error message gives a list of predicates that are not domain restricted.
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  weakly restricted variables: { ... }"
\end{verbatim}
This error message gives a list of variables that are not lambda restricted. 
\item
\begin{verbatim}
 "syntax error on line Y column Z unexpected token: X"
\end{verbatim}
The parser can not read the input format.
\newline
The next errors all belong to the ASPils output format.
Some constructs in the input language can not be expressed in certain normal forms.
Another normal form as to be chosen to support this feature.
\begin{verbatim}
 "Disjunction not allowed on this normal form,
  please choose normal form 2 or 4-7."
 "Count aggregate/cardinality constraint not allowed
  in this normal form,
  please choose normal form 3-7."
 "Count aggregate/cardinality constraint not allowed 
  with non trivial upper bound in this normal form,
  please choose normal form 5-7."
 "Count aggregate/cardinality constraint not allowed 
  with negative lower bound in this normal form,
  please choose normal form 5-7."
 "Sum aggregate/weight constraint not allowed
  in this normal form,
  please choose normal form 3-7."
 "Sum aggregate/weight constraint not allowed 
  with non trivial upper bound in this normal form,
  please choose normal form 5-7."
 "Sum aggregate/weight constraint not allowed 
  with negative lower bound in this normal form,
  please choose normal form 5-7"
 "Sum aggregate/weight constraint not allowed
  with negative weights in this normal form,
  please choose normal form 5 or 7."
 "Max aggregate not allowed in this normal form,
  please choose normal form 6 or 7."
 "Max aggregate not allowed
  with negative weights in this normal form,
  please choose normal form 7."
 "Min aggregate not allowed in this normal form,
  please choose normal form 6 or 7."
 "Min aggregate not allowed 
  with negative weights in this normal form,
  please choose normal form 7."
 "Optimize statement not allowed in this normal form,
  please choose normal form 3-7."
\end{verbatim}
%\begin{verbatim}
% "Times aggregate not allowed in this normal form."
% "Times aggregate not allowed
%  with negative weights in this normal form."
%\end{verbatim}
\end{itemize}
